---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  eval = TRUE
)

library(bbsTaiwan)
```

# bbsTaiwan

<!-- badges: start -->

[![R-CMD-check](https://github.com/SunnyTseng/bbsTaiwan/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/SunnyTseng/bbsTaiwan/actions/workflows/R-CMD-check.yaml)

<!-- badges: end -->

The goal of bbsTaiwan is to streamline Taiwan Breeding Birds Survey (BBS) data retrieval and analysis. It will support data retrieval from GBIF, where Taiwan BBS data are stored. ‘bbsTrim’ will also include the population trend modelling. Users can tailor species analysis, controlling data stratification, preparation, and modeling. The goal is to expand Taiwan BBS data access to the broader R-society in Taiwan.

## 📑 Background

## 💻 Set up

You can install and load the development version of bbsTaiwan from [GitHub](https://github.com/) with:

```{r eval = FALSE}
# install.packages("devtools")
devtools::install_github("SunnyTseng/bbsTaiwan")

# load the package
library(bbsTaiwan)
```

## ⛺ How to use

### If you need raw Taiwan BBS dataset on GBIF:

Taiwan BBS raw data can be accessed directly by typing the object name:

```{r eval = FALSE}
bbsTaiwan::occurrence
bbsTaiwan::event
bbsTaiwan::measurementorfacts
bbsTaiwan::extendedmeasurementorfact
```

The complete Taiwan BBS dataset on GBIF is following the Darwin Core Archive (DwC - A) format:

-   Occurrence: lists a set of times and locations at which particular species have been recorded

-   Event: lists includes the protocols used, the sample size, and the location for each

-   Measurement or Facts: includes additional information relating to the events

-   Extended Measurement or Fact: includes additional information relating to the taxon occurrences

### `bbsTaiwan` also provide a workflow to process Taiwan BBS data:

This is a basic example which shows you how to use the package:

1.  Check out the distribution of the BBS sites and the number of sites that have been surveyed throughout the years:

```{r bbs_history}
bbs_history()
```

1.  Look up the scientific name for species of interest

```{r bbs_translate}
bbs_translate(c("白頭翁", "烏頭翁", "台灣噪眉"))
```

2.  Fetch the data for the species of interest. If you need the dataset across all species, leave the argument `target_species` as `NULL`.

```{r bbs_fetch}
x <- bbs_fetch(target_species = bbs_translate("火冠戴菊鳥"))
```

The output from `bbs_fetch` is a list with two elements: `occurrence` and `site_info`. `occurrence` is a tibble, showing all the observations for the target species within selected year range

```{r bbs_fetch_out1}
x$occurrence
```

The second element `site_info` is a tibble including all the BBS sites:

```{r bbs_fetch_out2}
x$site_info
```

The function `bbs_fetch` can also take multiple species at the same time:

```{r bbs_fetch_out3}
x <- bbs_fetch(bbs_translate(c("烏頭翁", "白頭翁")))
```

3.  View the distribution of all the occurrence for the target species within selected year range

```{r bbs_plotmap}
bbs_plotmap(x)
```

4.  Get basic summary statistics for the fetched data

```{r bbs_stat}
bbs_stat(x)
```
